---
title: 'Lab: Class One Survey Analysis'
output:
  html_document:
---

### **INSTRUCTIONS**

#### **Complete all tasks 1-7 using reproducible coding practices (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks.**

**0. Install packages and open libraries.**

```{r}
install.packages("pacman")
pacman::p_load(haven, readr, dplyr, openxlsx)
```

##### 1. Import class one survey data from our Github site (<https://github.com/kijohnson/Advanced-Data-Analysis>) (The dataset is called 'Class 1 Survey Fall 2026.csv'), calling the R dataframe that you create *C1survey*.

```{r}
C1survey <-
  read_csv(
    "https://github.com/kijohnson/Advanced-Data-Analysis/raw/refs/heads/main/Class%201%20Survey%20Fall%202026.csv"
  )

```

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.

```{r}
dim(C1survey)
# a. 29 observations, b. 27 variables
```

##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed. **Hints**: You can use either the rename_with function or the names function. The list of names for time is:

```{r, echo=TRUE}
c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
```

```{r}
names(C1survey)[1:27] <- c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
```

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset. **Suggestion**: Try to ask ChatGPT for a solution if you are unsure. There is a really short answer to this but we want to see what students come up with!

```{r}
var_types <- sapply(C1survey, class)
table(var_types)
```

##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values (**hint**: use table function). If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. \*Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}
table(C1survey$bday, useNA = "always") 
table(C1survey$bmonth, useNA = "always") 

#I will remove the na/missing values
C1survey_NoNA <- na.omit(C1survey)

C1survey_NoNA$bday <- as.numeric(C1survey_NoNA$bday)
C1survey_NoNA$bmonth <- as.numeric(C1survey_NoNA$bmonth)

median(C1survey_NoNA$bday, na.rm=TRUE) 
median(C1survey_NoNA$bmonth, na.rm=TRUE) 
```

##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

```{r}
# (A) Convert birth month and day into a proper Date
# Create season variable
C1survey_NoNA$bseason <- with(C1survey_NoNA, ifelse(
  (bmonth == 12 & bday >= 1) | (bmonth == 1) | (bmonth == 2 & bday <= 29), "Winter",
  ifelse((bmonth >= 3 & bmonth <= 5), "Spring",
         ifelse((bmonth >= 6 & bmonth <= 8), "Summer",
                ifelse((bmonth >= 9 & bmonth <= 11), "Fall", NA)))))

# (B)
  table(C1survey_NoNA$bmonth, C1survey_NoNA$bseason, useNA = "always")
  
# (C)
  season_table <- table(C1survey_NoNA$bmonth, C1survey_NoNA$bseason)
  addmargins(season_table, margin = 2)
  
        
```

##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. \#'s).

```{r}

#Question: How many students said that their favorite day of the week is Tuesday? 

sum(C1survey_NoNA$fav_day == "Tuesday", na.rm = TRUE)

#Answer: 1 student
```


##### 8. 

```{r}

write.csv(C1survey_NoNA,
          "class1survey_cleaned.csv",
          row.names = FALSE)

```

